---
title: "RSVPLG02: Filter Data"
author: "David Fencsik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
this_file = "02-FilterData"
```

# Optionally Remove Participants

## Load data files

```{r, loaddata}
rsvpData = read.csv("rsvpData.csv")
surveyData = read.csv("surveyData.csv")
rsvpAggregate = read.csv("rsvpAggregate.csv")
```

## Number of Particants Before Filtering

```{r, preOutlierRemovalSummary}
print(length(unique(surveyData$sub)))
print(length(unique(rsvpAggregate$sub)))
```

## Remove participants with low accuracy

Participants 424 and 524 are barely above chance on T1 (T1 < 60%)

```{r, lowaccuracy}
runFilter = TRUE
if (runFilter) {
  subToRemove = c(424, 524)
  rsvpData = rsvpData |>
    filter(!(sub %in% subToRemove))
  surveyData = surveyData |>
    filter(!(sub %in% subToRemove))
  rsvpAggregate = rsvpAggregate |>
    filter(!(sub %in% subToRemove))
}
```

## Number of Particants After Filtering

```{r, postOutlierRemovalSummary}
print(length(unique(surveyData$sub)))
print(length(unique(rsvpAggregate$sub)))
```

## Save data

```{r, savedata}
save(rsvpData, surveyData, rsvpAggregate, file=sprintf("%s.RData", this_file))
```
