---
title: "RSVPLG02: Accuracy Analyses"
author: "David Fencsik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(rstatix)
library(ez)
this_file = "06-Accuracy"
source_file = "02-FilterData"
dataFile = sprintf("%s.RData", source_file)
```

# Load data files

```{r, loaddata1}
if(!file.exists(dataFile)) {
  sprintf("Cannot find file %s", dataFile)
  knitr::knit_exit()
}
```

```{r, loaddata2}
load(dataFile)
rsvpAggregate2 = rsvpAggregate
load("../rsvplg01/02-FilterData.RData")
rsvpAggregate1 = rsvpAggregate
rsvpAggregate = rsvpAggregate2
```

# Basic Accuracy Analyses

```{r basics1, fig.height=4, fig.width=4}
rsvpAggregate |>
  group_by(sub, bsam, sias6, promisa, promisd) |>
  summarize(t1acc = mean(t1acc),
            t2acc = mean(t2t1acc),
            .groups="drop") |>
  pivot_longer(cols = ends_with("acc"),
               names_to = "target",
               values_to = "acc") |>
  mutate(target = factor(target,
                         levels=c("t1acc", "t2acc"),
                         labels=c("T1", "T2"))) -> x
```

Violin plots of overall T1 and T2|T1 accuracy

```{r basics2, fig.height=4, fig.width=4}
x |>
  ggplot(aes(target, acc)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="Target", y="Accuracy") +
  theme_classic()
```

Plot accuracy in each task as a function of PROMIS-A

```{r basics3, fig.height=4, fig.width=6}
x |>
  ggplot(aes(x=promisa, y=acc, color=target)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  scale_color_manual(values=c("black","red")) +
  theme(legend.position="right",
        legend.title=element_blank()) +
  labs(x="Sum of PROMIS-A", y="Accuracy") +
  theme_classic()
```

Linear model of the effect of PROMIS-A on T1 accuracy

```{r basics4}
fit = lm(acc ~ promisa, data=x, subset=(target == "T1"))
summary(fit)
confint(fit, 'promisa', level=0.95)
```

Same but for T2 accuracy

```{r basics5}
fit = lm(acc ~ promisa, data=x, subset=(target == "T2"))
summary(fit)
confint(fit, 'promisa', level=0.95)
```


# Combine Data Sets w/ RSVPLG01

```{r, combinedata}
if (all(names(rsvpAggregate2) == names(rsvpAggregate1))) {
  rsvpAggregate2$experiment = "RSVPLG02"
  rsvpAggregate1$experiment = "RSVPLG01"
  combined = bind_rows(rsvpAggregate1,rsvpAggregate2)
} else {
  combined = NA
}
```

# RSVP Analyses

Generate standard RSVP plots looking at T1 and T2|T1 accuracy as a function of lag

## T1 and T2|T1 accuracy as a function of lag, collapsing across local/global
```{r, rsvpplots1, fig.height=4, fig.width=6}
combined |>
  drop_na(t2t1acc) |>
  filter(lag > 0) |>
  group_by(experiment, sub, lag) |>
  summarize(t1acc = mean(t1acc),
            t2acc = mean(t2t1acc),
            .groups="drop") |>
  group_by(experiment, lag) |>
  summarize(t1acc = mean(t1acc),
            t2acc = mean(t2acc),
            .groups="drop") |>
  pivot_longer(cols = ends_with("acc"),
               names_to = "target",
               values_to = "acc") |>
  mutate(target = factor(sub("acc", "", target)),
         experiment = factor(experiment, levels=c("RSVPLG02", "RSVPLG01")),
         lag = factor(lag)) |>
  ggplot(aes(x=lag, y=acc, color=target, linetype=experiment,
             group=interaction(target, experiment))) +
  geom_line(linewidth=1.25) +
  geom_point(size=3) +
  scale_color_manual(values=c("black","red")) +
  theme_classic() +
  theme(legend.position="right",
        legend.title=element_blank()) +
  coord_cartesian(ylim=c(0.5, 1.0))
#  facet_wrap(~lag)
```

Here is a basic plot of the AB just for this experiment

```{r abplot1, fig.height=4, fig.width=6}
x = rsvpAggregate |>
  group_by(sub, lag, bsam, sias6, promisa, promisd) |>
  summarize(t1acc = mean(t1acc),
            t2acc = mean(t2t1acc),
            .groups="drop")

x |>
  group_by(lag) |>
  summarize(t1acc = mean(t1acc),
            t2acc = mean(t2acc),
            .groups="drop") |>
  pivot_longer(cols = ends_with("acc"),
               names_to = "target",
               values_to = "acc") |>
  mutate(target = factor(sub("acc", "", target),
                         levels=c("t1", "t2"),
                         labels=c("T1", "T2|T1")),
         lag = factor(lag)) |>
  ggplot(aes(x=lag, y=acc, group=target, color=target)) +
  geom_line(linewidth=1.25) +
  geom_point(size=3) +
  scale_color_manual(values=c("black","red")) +
  coord_cartesian(ylim=c(0.5, 1.0)) +
  labs(x="T2 Lag (ms)", y="Accuracy") +
  theme(legend.position="right",
        legend.title=element_blank()) +
  theme_classic()
```

Plot T2|T1 accuracy as a function of lag, separating by each of the T1 and T2 levels

```{r abplotbylevels}
rsvpAggregate |>
  mutate(levels = sprintf("%s-%s", t1_level, t2_level)) |>
  group_by(sub, lag, levels) |>
  summarize(t2acc = mean(t2t1acc), .groups="drop") |>
  group_by(lag, levels) |>
  summarize(t2acc = mean(t2acc), .groups="drop") |>
  ggplot(aes(x=lag, y=t2acc, group=levels, color=levels)) +
  geom_line(linewidth=1.25) +
  geom_point(size=3) +
#  scale_color_manual(values=c("black","red")) +
  coord_cartesian(ylim=c(0.5, 1.0)) +
  labs(x="T2 Lag (ms)", y="T2|T1 Accuracy") +
  theme(legend.position="right",
        legend.title=element_blank()) +
  theme_classic()
```

ANOVA testing the effects of lag and combined levels on T2|T1 accuracy

```{r}
rsvpAggregate |>
  mutate(levels = sprintf("%s-%s", t1_level, t2_level)) |>
  group_by(sub, lag, levels) |>
  summarize(t2acc = mean(t2t1acc), .groups="drop") |>
  anova_test(dv=t2acc,
             wid=sub,
             within=c(lag, levels))
```


Violin plots of T1 accuracy separated by lag

```{r abplott1, fig.height=4, fig.width=4}
x |>
  mutate(lag = factor(lag)) |>
  ggplot(aes(lag, t1acc)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="Lag", y="T1 Accuracy") +
  theme_classic()
```

Paired t-test comparing T1 accuracy across the two lags, just for kicks. If it's significant, presumably that's just because we have so many participants?

```{r abtest1}
x |>
  t_test(t1acc ~ lag, paired=TRUE)
```

Violin plots of T2|T1 accuracy separated by lag

```{r abplot2, fig.height=4, fig.width=4}
x |>
  mutate(lag = factor(lag)) |>
  ggplot(aes(lag, t2acc)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="Lag", y="T2|T1 Accuracy") +
  theme_classic()
```

Paired t-test comparing T2|T1 accuracy across the two lags

```{r abtest2}
x |>
  t_test(t2acc ~ lag, paired=TRUE)
```

Plots of difference scores based on T2|T1 accuracy lag 750 - lag 300

```{r abplot3, fig.height=4, fig.width=3}
diff = x |>
  select(!t1acc) |>
  mutate(lag = sprintf("l%d", lag)) |>
  pivot_wider(names_from=lag,
              values_from=t2acc) |>
  mutate(d = l750 - l300)

diff |>
  ggplot(aes("", d)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  geom_hline(yintercept=0, linetype=2, color="gray") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x=NULL, y="Lag 750 - Lag 300") +
  theme_classic()
```

Plot the difference score as a function of PROMIS-A

```{r abplot4, fig.height=4, fig.width=6}
diff |>
  ggplot(aes(x=promisa, y=d)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
#  scale_color_manual(values=c("black","red")) +
  theme(legend.position="right",
        legend.title=element_blank()) +
  labs(x="Sum of PROMIS-A", y="Lag 750 - Lag 300") +
  theme_classic()
```

Run a linear model on the above function

```{r abplot4lm}
fit = lm(d ~ promisa, data=diff)
summary(fit)
confint(fit, 'promisa', level=0.95)
```


## Individual accuracy

Violin plot of overall accuracy with each target

```{r, indaccplot, fig.height=4, fig.width=4}
x = rsvpAggregate2 |>
  drop_na(t2t1acc) |>
  group_by(sub) |>
  summarize(t1acc = mean(t1acc),
            t2acc = mean(t2t1acc),
            .groups="drop")
x |>
  pivot_longer(cols = ends_with("acc"),
               names_to = "target",
               values_to = "acc") |>
  mutate(target = factor(sub("acc", "", target))) |>
  ggplot(aes(target, acc)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  theme_classic() +
  theme(legend.position="none") +
  scale_x_discrete("Target",
    labels = c("t1"="T1", "t2"="T2|T1")) +
  coord_cartesian(ylim=c(0.5, 1.0))
```

Participants with the lowest T1 accuracy

```{r, indacclist1}
x |>
  arrange(t1acc) |>
  head(10)
```

Participants with the lowest T2 accuracy

```{r, indacclist2}
x |>
  arrange(t2acc) |>
  head(10)
```

# Effects of anxiety on AB and local/global processing

Create groups

```{r creategroups}
nGroups = 2
dt = rsvpAggregate2 |>
  mutate(anx = promisa) |>
  drop_na(anx) |>
  mutate(anxgrp = cut(anx, nGroups, labels=as.character(1:nGroups)))
```

Test grouping function

```{r testgroups}
dt |>
  ggplot(aes(anxgrp, anx)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  theme(legend.position="none") +
  theme_classic()
```

Plot by groups

```{r plotgroupsandscope}
dt = dt |>
  filter(anxgrp == "1" | anxgrp == as.character(nGroups)) |>
  mutate(anxgrp = factor(anxgrp),
         scope = paste(t1_level, t2_level, sep="-"))

dt |>
  group_by(scope, lag, anxgrp) |>
  summarize(t2acc = mean(t2t1acc), .groups="drop") |>
  ggplot(aes(lag, t2acc, color=anxgrp)) +
  geom_line(linewidth=1.25) +
  facet_wrap(~scope) +
  theme_classic()
```

```{r anovagroupsandscope}
# get the 4 combos of scope
dt = dt |>
  mutate(sub=factor(sub),
         lag=factor(lag))

scope = sort(unique(dt$scope))

for (s in scope) {
  cat("ANOVA on scope", s, "\n")
  # print(summary(dt[dt$scope == s, ]))
  dt |>
    filter(scope == s) |>
    ezANOVA(dv=t2t1acc,
                wid=sub,
                within=lag,
                between=anxgrp,
                type=3,
                detailed=FALSE) |>
    print()
}
```
