---
title: "RSVPLG02: Accuracy Analyses"
author: "David Fencsik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(rstatix)
library(ez)
this_file = "06-Accuracy"
source_file = "02-FilterData"
dataFile = sprintf("%s.RData", source_file)
```

# Load data files

```{r, loaddata1}
if(!file.exists(dataFile)) {
  sprintf("Cannot find file %s", dataFile)
  knitr::knit_exit()
}
```

```{r, loaddata2}
load(dataFile)
rsvpAggregate2 = rsvpAggregate
load("../rsvplg01/02-FilterData.RData")
rsvpAggregate1 = rsvpAggregate
```

# Combine Data Sets

```{r, combinedata}
if (all(names(rsvpAggregate2) == names(rsvpAggregate1))) {
  rsvpAggregate2$experiment = "RSVPLG02"
  rsvpAggregate1$experiment = "RSVPLG01"
  combined = bind_rows(rsvpAggregate1,rsvpAggregate2)
} else {
  combined = NA
}
```

# RSVP Analyses

Generate standard RSVP plots looking at T1 and T2|T1 accuracy as a function of lag

## T1 and T2|T1 accuracy as a function of lag, collapsing across local/global
```{r, rsvpplots1, fig.height=4, fig.width=6}
combined |>
  drop_na(t2t1acc) |>
  filter(lag > 0) |>
  group_by(experiment, sub, lag) |>
  summarize(t1acc = mean(t1acc),
            t2acc = mean(t2t1acc),
            .groups="drop") |>
  group_by(experiment, lag) |>
  summarize(t1acc = mean(t1acc),
            t2acc = mean(t2acc),
            .groups="drop") |>
  pivot_longer(cols = ends_with("acc"),
               names_to = "target",
               values_to = "acc") |>
  mutate(target = factor(sub("acc", "", target)),
         experiment = factor(experiment, levels=c("RSVPLG02", "RSVPLG01")),
         lag = factor(lag)) |>
  ggplot(aes(x=lag, y=acc, color=target, linetype=experiment,
             group=interaction(target, experiment))) +
  geom_line(linewidth=1.25) +
  geom_point(size=3) +
  scale_color_manual(values=c("black","red")) +
  theme_classic() +
  theme(legend.position="right",
        legend.title=element_blank()) +
  coord_cartesian(ylim=c(0.5, 1.0))
#  facet_wrap(~lag)
```

## Individual accuracy

Violin plot of overall accuracy with each target

```{r, indaccplot, fig.height=4, fig.width=4}
x = rsvpAggregate2 |>
  drop_na(t2t1acc) |>
  group_by(sub) |>
  summarize(t1acc = mean(t1acc),
            t2acc = mean(t2t1acc),
            .groups="drop")
x |>
  pivot_longer(cols = ends_with("acc"),
               names_to = "target",
               values_to = "acc") |>
  mutate(target = factor(sub("acc", "", target))) |>
  ggplot(aes(target, acc)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  theme_classic() +
  theme(legend.position="none") +
  scale_x_discrete("Target",
    labels = c("t1"="T1", "t2"="T2|T1")) +
  coord_cartesian(ylim=c(0.5, 1.0))
```

Participants with the lowest T1 accuracy

```{r, indacclist1}
x |>
  arrange(t1acc) |>
  head(10)
```

Participants with the lowest T2 accuracy

```{r, indacclist2}
x |>
  arrange(t2acc) |>
  head(10)
```

# Effects of anxiety on AB and local/global processing

Create groups

```{r creategroups}
nGroups = 2
dt = rsvpAggregate2 |>
  mutate(anx = promisa) |>
  drop_na(anx) |>
  mutate(anxgrp = cut(anx, nGroups, labels=as.character(1:nGroups)))
```

Test grouping function

```{r testgroups}
dt |>
  ggplot(aes(anxgrp, anx)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  theme(legend.position="none") +
  theme_classic()
```

Plot by groups

```{r plotgroupsandscope}
dt = dt |>
  filter(anxgrp == "1" | anxgrp == as.character(nGroups)) |>
  mutate(anxgrp = factor(anxgrp),
         scope = paste(t1_level, t2_level, sep="-"))

dt |>
  group_by(scope, lag, anxgrp) |>
  summarize(t2acc = mean(t2t1acc), .groups="drop") |>
  ggplot(aes(lag, t2acc, color=anxgrp)) +
  geom_line(linewidth=1.25) +
  facet_wrap(~scope) +
  theme_classic()
```

```{r anovagroupsandscope}
# get the 4 combos of scope
dt = dt |>
  mutate(sub=factor(sub),
         lag=factor(lag))

scope = sort(unique(dt$scope))

for (s in scope) {
  cat("ANOVA on scope", s, "\n")
  # print(summary(dt[dt$scope == s, ]))
  dt |>
    filter(scope == s) |>
    ezANOVA(dv=t2t1acc,
                wid=sub,
                within=lag,
                between=anxgrp,
                type=3,
                detailed=FALSE) |>
    print()
}
```
