---
title: "RSVPLG02: Switch-Cost Analyses"
author: "David Fencsik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggpubr)
library(rstatix)
this_file = "08-Switching"
source_file = "02-FilterData"
dataFile = sprintf("%s.RData", source_file)
```

# Summary

The purpose of this analysis is to look at switch costs, starting with whether there is any evidence of switching between scopes.

# Load data files

```{r, loaddata1, include=FALSE}
if(!file.exists(dataFile)) {
  sprintf("Cannot find file %s", dataFile)
  knitr::knit_exit()
}
```

```{r, loaddata2}
load(dataFile)
```

# Global/Local Differences

## T1

```{r globallocal1}
rsvpAggregate |>
  group_by(sub, t1_level, bsam, sias6, promisa, promisd) |>
  summarize(acc = mean(t1acc), .groups="drop") |>
  rename(level = t1_level) -> x
```

T1 accuracy as a function of T1 level

```{r globallocal2, fig.height=4, fig.width=4}
x |>
  ggplot(aes(level, acc)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="T1 Level", y="T1 Accuracy") +
  theme_classic()
```

T1 accuracy as a function of T1 level and PROMIS-A score

```{r globallocal3, fig.height=4, fig.width=6}
x |>
  ggplot(aes(x=promisa, y=acc, color=level)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  scale_color_manual(values=c("black","red")) +
  theme(legend.position="right",
        legend.title=element_blank()) +
  labs(x="Sum of PROMIS-A", y="T1 Accuracy") +
  theme_classic()
```

Linear fit for T1 global level

```{r globallocal3lm1}
fit = lm(acc ~ promisa, data=x, subset=(level == "global"))
summary(fit)
confint(fit, 'promisa', level=0.95)
```

Linear fit for T1 local level

```{r globallocal3lm2}
fit = lm(acc ~ promisa, data=x, subset=(level == "local"))
summary(fit)
confint(fit, 'promisa', level=0.95)
```


## T2

```{r globallocal4}
rsvpAggregate |>
  group_by(sub, t2_level, bsam, sias6, promisa, promisd) |>
  summarize(acc = mean(t2acc), .groups="drop") |>
  rename(level = t2_level) -> x
```

T2 accuracy as a function of T2 level

```{r globallocal5, fig.height=4, fig.width=4}
x |>
  ggplot(aes(level, acc)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="T2 Level", y="T2|T1 Accuracy") +
  theme_classic()
```

T2 accuracy as a function of T2 level and PROMIS-A score

```{r globallocal6, fig.height=4, fig.width=6}
x |>
  ggplot(aes(x=promisa, y=acc, color=level)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  scale_color_manual(values=c("black","red")) +
  theme(legend.position="right",
        legend.title=element_blank()) +
  labs(x="Sum of PROMIS-A", y="T2|T1 Accuracy") +
  theme_classic()
```

Linear fit for T2 global level

```{r globallocal6lm1}
fit = lm(acc ~ promisa, data=x, subset=(level == "global"))
summary(fit)
confint(fit, 'promisa', level=0.95)
```

Linear fit for T2 local level

```{r globallocal6lm2}
fit = lm(acc ~ promisa, data=x, subset=(level == "local"))
summary(fit)
confint(fit, 'promisa', level=0.95)
```



# Overall Switching Effects

```{r overallswitch1}
x = rsvpAggregate |>
  mutate(switch = ifelse(t1_level == t2_level, "noswitch", "switch")) |>
  group_by(sub, switch, bsam, sias6, promisa, promisd) |>
  summarize(acc = mean(t2t1acc), .groups="drop")

diff = x |>
  pivot_wider(names_from=switch,
              values_from=acc) |>
  mutate(d = noswitch - switch)
```

Plot of T2|T1 accuracy on switch and no switch trials

```{r overallswitch2, fig.height=4, fig.width=4}
x |>
  ggplot(aes(switch, acc)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="Switch Condition", y="T2|T1 Accuracy") +
  scale_x_discrete(labels=c("No Switch", "Switch")) +
  theme_classic()
```

Plot of difference score: T2|T1 accuracy, no-switch minus switch

```{r overallswitch3, fig.height=4, fig.width=4}
diff |>
  ggplot(aes(x="", y=d)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_hline(yintercept=0, linetype=2, color="gray") +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x=NULL, y="No-Switch - Switch") +
  theme_classic()
```

Summary statistics for T2|T1 accuracy on no-switch and switch trials

```{r overallswitch4}
x |>
  group_by(switch) |>
  get_summary_stats(acc, type = "mean_sd")
```

Shapiro-Wilk test of normality for the difference score. A significant outcome indicates non-normality

```{r overallswitch5}
diff |>
  shapiro_test(d)
```

QQ plot of the difference score to test for non-normality

```{r overallswitch6, fig.height=4, fig.width=4}
ggqqplot(diff, "d")
```

Paired *t*-test comparing T2|T1 accuracy on no-switch and switch trials

```{r overallswitch7}
x |>
  t_test(acc ~ switch, paired=TRUE)
```

## Effect of Surveys

Plot the difference score as a function of PROMIS-A

```{r overallswitchbypromisa1, fig.height=4, fig.width=4}
diff |>
  ggplot(aes(x=promisa, y=d)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  theme(legend.position="right",
        legend.title=element_blank()) +
  labs(x="Sum of PROMIS-A", y="No-Switch - Switch") +
  theme_classic()

fit = lm(d ~ promisa, data=diff)
summary(fit)
confint(fit, 'promisa', level=0.95)
```


# Switch Costs by Level

```{r switchbylevel1}
x = rsvpAggregate |>
  mutate(switch = ifelse(t1_level == t2_level, "noswitch", "switch")) |>
  group_by(sub, switch, t2_level, bsam, sias6, promisa, promisd) |>
  summarize(acc = mean(t2t1acc), .groups="drop")

diff = x |>
  pivot_wider(names_from=switch,
              values_from=acc) |>
  mutate(d = noswitch - switch)
```

Violin plot of T2|T1 accuracy separately for each combination of T2 level and switch condition

```{r switchbylevel2, fig.height=4, fig.width=8}
x |>
  ggplot(aes(switch, acc)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="Switch Condition", y="T2|T1 Accuracy") +
  scale_x_discrete(labels=c("No Switch", "Switch")) +
  facet_wrap(~t2_level, labeller=as_labeller(c(local="T2 Local", global="T2 Global"))) +
  theme_classic()
```

Violin plot of the difference score for T2 local or global. The difference score is accuracy on no-switch trials minus accuracy on switch trials

```{r switchbylevel3, fig.height=4, fig.width=4}
diff |>
  ggplot(aes(x=t2_level, y=d)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_hline(yintercept=0, linetype=2, color="gray") +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="T2 Level", y="No-Switch - Switch") +
  scale_x_discrete(labels=c("Global", "Local")) +
  theme_classic()
```

Summary statistics for T2|T1 accuracy for each of the 4 combinations of switch-type and T2 level

```{r switchbylevel4}
x |>
  group_by(t2_level, switch) |>
  get_summary_stats(acc, type = "mean_sd")
```

Summary statistics for the difference score for each T2 level

```{r switchbylevel4a}
diff |>
  group_by(t2_level) |>
  get_summary_stats(d, type = "mean_sd")
```

Shapiro-Wilk test of normality for accuracy. A significant outcome supports non-normality

```{r switchbylevel5a}
x |>
  group_by(t2_level, switch) |>
  shapiro_test(acc)
```

Shapiro-Wilk test of normality for the difference score. A significant outcome supports non-normality

```{r switchbylevel5b}
diff |>
  group_by(t2_level) |>
  shapiro_test(d)
```

QQ plot of the difference scores at each T2 level

```{r switchbylevel6, fig.height=4, fig.width=8}
diff |>
  ggqqplot("d", facet.by="t2_level")
```

Overall repeated-measures ANOVA looking at the effects of T2 level (global, local) and switch condition (no-switch, switch) on T2|T1 accuracy

```{r switchbylevel7}
x |>
  anova_test(dv=acc,
             wid=sub,
             within=c(t2_level, switch))
```

Planned comparisons for the above ANOVA: paired *t*-tests looking at the switch cost for each T2 level

```{r switchbylevel8}
x |>
  group_by(t2_level) |>
  t_test(acc ~ switch, paired=TRUE)
```

Now run a *t*-tests on the difference scores for each T2 level. These are more appropriate because the difference scores are normally distributed, whereas some of the accuracies don't seem to be.

Run at *t*-test comparing the difference scores at each level. That is, compare the switch cost at T2 global to the switch cost at T2 local

```{r switchbylevel9}
diff |>
  t_test(d ~ t2_level, paired=TRUE)
```

# Difference Scores by Level and by Surveys

Plot local and global difference scores by PROMIS-A

```{r diffbypromisa1}
diff |>
  ggplot(aes(x=promisa, y=d, color=t2_level)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  scale_color_manual(values=c("black","red")) +
  theme(legend.position="right",
        legend.title=element_blank()) +
  labs(x="Sum of PROMIS-A", y="No-Switch - Switch") +
  theme_classic()
```

Separate plots for global and local, but otherwise same as above

```{r diffbypromisa2, fig.height=4, fig.width=8}
diff |>
  ggplot(aes(x=promisa, y=d)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  facet_wrap(~t2_level, labeller=as_labeller(c(local="T2 Local", global="T2 Global"))) +
  theme(legend.position="right",
        legend.title=element_blank()) +
  labs(x="Sum of PROMIS-A", y="No-Switch - Switch") +
  theme_classic()
```

Linear fit of global data above. Runs a regression of the effect of PROMIS-A on the difference score for just T2 global.

```{r diffbypromisalm1}
fit = lm(d ~ promisa, data=diff, subset=(t2_level == "global"))
summary(fit)
confint(fit, 'promisa', level=0.95)
```

Same as above but for T2 local.

```{r diffbypromisalm2}
fit = lm(d ~ promisa, data=diff, subset=(t2_level == "local"))
summary(fit)
confint(fit, 'promisa', level=0.95)
```


# Overall Switch Costs by Lag

```{r switchbylag1}
x = rsvpAggregate |>
  mutate(switch = ifelse(t1_level == t2_level, "noswitch", "switch"),
         lag = factor(lag)) |>
  group_by(sub, switch, lag, bsam, sias6, promisa, promisd) |>
  summarize(acc = mean(t2t1acc), .groups="drop")

diff = x |>
  pivot_wider(names_from=switch,
              values_from=acc) |>
  mutate(d = noswitch - switch)
```

Violin plot of T2|T1 accuracy separately for each combination of T2 lag and switch condition

```{r switchbylag2, fig.height=4, fig.width=8}
x |>
  ggplot(aes(switch, acc)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="Switch Condition", y="T2|T1 Accuracy") +
  scale_x_discrete(labels=c("No Switch", "Switch")) +
  facet_wrap(~lag) +
  theme_classic()
```

Violin plot of the difference score for each level of T2 lag. The difference score is still accuracy on no-switch trials minus accuracy on switch trials

```{r switchbylag3, fig.height=4, fig.width=4}
diff |>
  ggplot(aes(x=lag, y=d)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_hline(yintercept=0, linetype=2, color="gray") +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="T2 Lag (ms)", y="No-Switch - Switch") +
  theme_classic()
```

Overall repeated-measures ANOVA looking at the effects of T2 lag (300 vs. 750 ms) and switch condition (no-switch, switch) on T2|T1 accuracy. Any effect of T2 lag would be the attentional blink.

```{r switchbylag4}
x |>
  anova_test(dv=acc,
             wid=sub,
             within=c(lag, switch))
```

Planned comparisons for the above ANOVA: paired *t*-tests looking at the switch cost for each level of T2 lag

```{r switchbylag5}
x |>
  group_by(lag) |>
  t_test(acc ~ switch, paired=TRUE)
```

Now run a *t*-tests on the difference scores for each level of T2 lag. These are more appropriate because the difference scores are normally distributed, whereas some of the accuracies don't seem to be.

Run at *t*-test comparing the difference scores at each level. That is, compare the switch cost at 300 ms lags to 750 ms lags

```{r switchbylag6}
diff |>
  t_test(d ~ lag, paired=TRUE)
```

# Effects of Surveys by Lag

Plot difference scores by PROMIS-A separately for each lag

```{r diffbylagbypromisa1}
diff |>
  ggplot(aes(x=promisa, y=d, color=lag)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  scale_color_manual(values=c("black","red")) +
  theme(legend.position="right",
        legend.title=element_blank()) +
  labs(x="Sum of PROMIS-A", y="No-Switch - Switch") +
  theme_classic()
```

Separate plots for each lag, but otherwise same as above

```{r diffbylagbypromisa2, fig.height=4, fig.width=8}
diff |>
  ggplot(aes(x=promisa, y=d)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  facet_wrap(~lag) +
  theme(legend.position="right",
        legend.title=element_blank()) +
  labs(x="Sum of PROMIS-A", y="No-Switch - Switch") +
  theme_classic()
```

Linear fit of data above. Runs a regression of the effect of PROMIS-A on the difference score for lag 300.

```{r diffbylagbypromisalm1}
fit = lm(d ~ promisa, data=diff, subset=(lag == "300"))
summary(fit)
confint(fit, 'promisa', level=0.95)
```

Same as above but for lag 750.

```{r diffbylagbypromisalm2}
fit = lm(d ~ promisa, data=diff, subset=(lag == "750"))
summary(fit)
confint(fit, 'promisa', level=0.95)
```
