---
title: "RSVPLG02: Expansion/Contraction"
author: "David Fencsik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(rstatix)
this_file = "09-ExpCon"
source_file = "02-FilterData"
dataFile = sprintf("%s.RData", source_file)
```

# Load data

```{r, loaddata2, echo=TRUE}
load(dataFile)
```

# Summary

The purpose is to look at effects of anxiety as a function of expansion and contraction, both from the Kosslyn perspective, which defines expansion as L-L to L-G and contraction as G-G to G-L, and my own idea of comparing L-G to G-L. The issue with the first one is that we end up comparing T2 local to T2 global, and those might not be the same. The 2nd idea ends up looking at relative performance as a function of anxiety so feels more straightforward to me. Maybe it's all the same thing?

# Compare T2 local to T2 global

```{r comparet2localglobal1}
rsvpAggregate |>
  group_by(sub, t2_level, bsam, sias6, promisa, promisd) |>
  mutate(t2_level = factor(t2_level)) |>
  summarize(acc = mean(t2t1acc), .groups="drop") -> acc.by.level
```

Violin plot comparing overall T2|T1 accuracy separated by T2 level

```{r comparet2localglobal2, fig.height=4, fig.width=4}
acc.by.level |>
  ggplot(aes(t2_level, acc)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="T2 Level", y="T2|T1 Accuracy") +
#  scale_x_discrete(labels=c("No Switch", "Switch")) +
  theme_classic()
```

## Effects of Anxiety

Scatter plot indicating effect of PROMIS-A total on T2|T1 accuracy, separated by T2 level

```{r comparet2localglobal3, fig.height=5, fig.width=6}
acc.by.level |>
  ggplot(aes(x=promisa, y=acc, color=t2_level)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  scale_color_manual(values=c("black","red")) +
  theme(legend.position="right",
        legend.title=element_blank()) +
  labs(x="Sum of PROMIS-A", y="T2|T1 Accuracy") +
#  facet_wrap(~t2_level) +
  theme_classic()
```

Linear model of the above plot for just global

```{r comparet2localglobal4}
fit = lm(acc ~ promisa, data=acc.by.level, subset=(t2_level == "global"))
summary(fit)
confint(fit, 'promisa', level=0.95)

```

Linear model of the above plot for just local

```{r comparet2localglobal5}
fit = lm(acc ~ promisa, data=acc.by.level, subset=(t2_level == "local"))
summary(fit)
confint(fit, 'promisa', level=0.95)
```


# Expansion and Contraction

Given that T2 accuracy seems to be substantially different depending on whether T2 was global or local, I feel like Kosslyn's comparisons are a bit odd. I'll try the simpler version of them, defining expansion as L-G and contraction and G-L, ignoring trials with no switch

```{r expansioncontraction1}
rsvpAggregate |>
  filter(t1_level != t2_level) |>
  mutate(switch_type = factor(ifelse(t1_level == "global", "contract", "expand"))) |>
  group_by(sub, switch_type, bsam, sias6, promisa, promisd) |>
  summarize(acc = mean(t2t1acc), .groups="drop") -> expand.contract
```

Violin plot of T2 accuracy separated by contraction and expansion

```{r expansioncontraction2, fig.height=4, fig.width=4}
expand.contract |>
  ggplot(aes(switch_type, acc)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="Switch Type", y="T2|T1 Accuracy") +
  scale_x_discrete(labels=c("Contraction", "Expansion")) +
  theme_classic()
```


```{r expansioncontractiondiff1}
expand.contract |>
  pivot_wider(names_from=switch_type,
              values_from=acc) |>
  mutate(d = contract - expand) -> diff
```

Violin plot of difference score

```{r expansioncontractiondiff2, fig.height=4, fig.width=4}
diff |>
  ggplot(aes(x="", y=d)) +
  geom_violin(draw_quantiles=c(.25, .5, .75)) +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x=NULL, y="T2|T1 Accuracy (Contraction - Exponsion)") +
  theme_classic()
```

Effect of anxiety on the difference between contraction and expansion

```{r expansioncontractiondiff3, fig.height=4, fig.width=4}
diff |>
  ggplot(aes(x=promisa, y=d)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  theme(legend.position="right",
        legend.title=element_blank()) +
  labs(x="Sum of PROMIS-A", y="T2|T1 Accuracy (Contraction - Exponsion)") +
  theme_classic()
```

```{r expansioncontractiondiff4}
fit = lm(d ~ promisa, data=diff)
summary(fit)
confint(fit, 'promisa', level=0.95)
```
