---
title: 'RSVPLG02: Modeling'
author: "David E. Fencsik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, cache=TRUE)
library(tidyverse)
library(rstatix)
library(lme4)
library(lmerTest)
theme_set(theme_classic())
this_file = "12-Catwalk"
source_file = "02-FilterData"
dataFile = sprintf("%s.RData", source_file)
load(dataFile)
```

# Kitchen Sink Model

## Prepare data

```{r extractdata1}
rsvpData |>
  filter(trial_type == "exp",
         t1_acc >= 0,
         t2_acc >= 0) |>
  mutate(t1_level=str_to_title(t1_level),
         t2_level=str_to_title(t2_level),
         switch_type=factor(case_when(
           t1_level == t2_level ~ "Repeat",
           t1_level != t2_level ~ "Switch",
           .default = NA)),
         t2_level=factor(t2_level),
         lag = factor(lag)) |>
  filter(t1_acc == 1) |>
  group_by(sub, t2_level, switch_type, lag) |>
  summarize(acc = mean(t2_acc),
            .groups="drop") -> x

surveyData |>
  select(sub,
         bsam=sumbsam, sias6=sumsias6,
         promisd=sumpromisd, promisa=sumpromisa) |>
  right_join(x,
             by=join_by(sub),
             unmatched="error",
             relationship="one-to-many") -> x

summary(x)
```

## Plot overall data

Plot experimental variables

```{r lineplots, fig.height=4, fig.width=8}
x |>
  group_by(switch_type, t2_level, lag) |>
  summarize(acc = mean(acc), .groups="drop") |>
  ggplot(aes(x=switch_type, y=acc, group=t2_level, color = t2_level)) +
  geom_line(linewidth=1.25) +
  geom_point(size=3, shape=21, fill="white", stroke=2) +
  labs(x="Switch Condition", y="T2|T1 Accuracy") +
  theme(legend.position="inside",
        legend.position.inside=c(0.55, 0.2),
        legend.title=element_blank()) +
  facet_wrap(~lag)
```

Plot accuracy as a function of trait anxiety (PROMIS-A total) for each of the 4 conditions separately by the two lags

```{r scatterplots, fig.height=4, fig.width=8}
x |>
  group_by(switch_type, t2_level, lag, promisa) |>
  summarize(acc = mean(acc), .groups="drop") |>
  ggplot(aes(x=promisa, y=acc, group=interaction(t2_level, switch_type),
             color=t2_level, linetype=switch_type)) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  labs(x="Anxiety", y="T2|T1 Accuracy") +
  guides(linetype=guide_legend(override.aes=list(color="black"))) +
  theme(legend.title=element_blank()) +
  # theme(legend.position="inside",
  #       legend.position.inside=c(0.55, 0.2),
  #       legend.title=element_blank()) +
  facet_wrap(~lag)
```

# Modeling

## Kitchen Sink

Run the kitchen sink model with all four variables---switch type, T2 level, lag, and anxiety---as fixed effects.

```{r kitchensink}
options(width=150)
model_ks = lmer(acc ~ switch_type * t2_level * lag * scale(promisa) + (1 | sub),
                data=x)
summary(model_ks)
```

I need to think about what the interactions with lag mean, but for now I want to move on to focusing on the questions we actually want to ask

## Focus Switching Effects

```{r focusswitchingmodels}
# basic model
options(width=150)
model1 = lmer(acc ~ switch_type * t2_level * scale(promisa) + (1 | sub),
              data=x)
summary(model1)

# add variable slope for switch_type
model2 = lmer(acc ~ switch_type * t2_level * scale(promisa) + (1 + switch_type | sub),
              data=x)
summary(model2) # model2 is a better fit

# compare model1 and model2: reliable so keep variable switch_type
anova(model1, model2)

# remove switch_type and add variable slope for t2_level, which makes the model singular
model3 = lmer(acc ~ switch_type * t2_level * scale(promisa) + (1 + t2_level | sub),
              data=x)
summary(model3)

anova(model2, model3)
```

```{r fitsateachlevel}
options(width=150)

# Run the models I want to run looking at effect of switch on
# anxiety effect separately for each level

# global
model_global1 = lmer(acc ~ switch_type * scale(promisa) + (1 | sub),
                     data=x, subset=t2_level=="Global")
summary(model_global1)
model_global2 = lmer(acc ~ switch_type * scale(promisa) + (1 + switch_type | sub),
                     data=x, subset=t2_level=="Global")
summary(model_global2)
anova(model_global1, model_global2) # 1 is better

# local
model_local1 = lmer(acc ~ switch_type * scale(promisa) + (1 | sub),
                    data=x, subset=t2_level=="Local")
summary(model_local1)
model_local2 = lmer(acc ~ switch_type * scale(promisa) + (1 + switch_type | sub),
                    data=x, subset=t2_level=="Local")
summary(model_local2)
anova(model_local1, model_local2) # 2 is better


```