---
title: "RSVPLG02: Prepare Data for Classes"
author: "David Fencsik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rstatix)
```

# Load data files

```{r, loaddata}
rsvpData = read.csv("rsvpData.csv")
surveyData = read.csv("surveyData.csv")
```

# Filter participants w/ low accuracy

```{r filterdata}
rsvpData |>
  filter(trial_type == "exp",
         t1_acc >= 0, t2_acc >= 0) |>
  group_by(sub) |>
  summarize(t1 = mean(t1_acc),
            t2 = mean(t2_acc)) |>
  filter(t1 > .6 & t2 > .6) |>
  pull(sub) -> sub.to.keep

rsvpData |> filter(sub %in% sub.to.keep) -> rsvpData
surveyData |> filter(sub %in% sub.to.keep) -> surveyData
```

# Modify variables

```{r modifyvars}
rsvpData |>
  filter(trial_type == "exp",
         t1_acc >= 0,
         t2_acc >= 0) |>
  mutate(blocktyp = factor(blocktyp),
         trial_type = factor(trial_type),
         t1_level = factor(substr(t1_level, 1, 1)),
         t2_level = factor(substr(t2_level, 1, 1)),
         switch_type = factor(
           if_else(t1_level == t2_level, "repeat", "switch"))) -> rsvpData
```

# Create wide data set

Generate a wide data set for import into JASP that has columns for T2|T1 accuracy for each of the 8 cells in the factorial table of T1 level (local, global), T2 level (local, global), and T2 lag (300, 750)

```{r generateT2}
rsvpData |>
  filter(t1_acc == 1) |>
  group_by(sub, t1_level, t2_level, lag) |>
  summarize(t2acc = mean(t2_acc),
            .groups = "drop") |>
  pivot_wider(names_from = c(t1_level, t2_level, lag),
              values_from = t2acc,
              names_prefix = "t2acc_") -> jasp_export
```

Add columns that separate T2|T1 accuracy by switch type

```{r separateswitch}
rsvpData |>
  filter(t1_acc == 1) |>
  group_by(sub, switch_type) |>
  summarize(t2acc = mean(t2_acc), .groups="drop") |>
  pivot_wider(names_from = switch_type,
              values_from = t2acc,
              names_prefix = "t2acc_") |>
  right_join(jasp_export,
             by=join_by(sub),
             unmatched="error",
             relationship="one-to-one") -> jasp_export
```

Add columns that separate T2|T1 accuracy by lag without scope

```{r separatelag}
rsvpData |>
  filter(t1_acc == 1) |>
  group_by(sub, lag) |>
  summarize(t2acc = mean(t2_acc), .groups="drop") |>
  pivot_wider(names_from = lag,
              values_from = t2acc,
              names_prefix = "t2acc_") |>
  right_join(jasp_export,
             by=join_by(sub),
             unmatched="error",
             relationship="one-to-one") -> jasp_export
```

Add columns that separate T2|T1 accuracy by T2 level

```{r separatet2level}
rsvpData |>
  filter(t1_acc == 1) |>
  group_by(sub, t2_level) |>
  summarize(t2acc = mean(t2_acc), .groups="drop") |>
  pivot_wider(names_from = c(t2_level),
              values_from = t2acc,
              names_prefix = "t2acc_t2lvl_") |>
  right_join(jasp_export,
             by=join_by(sub),
             unmatched="error",
             relationship="one-to-one") -> jasp_export
```

Add columns that separate T2|T1 accuracy by scope global/local without lag

```{r separatelevels}
rsvpData |>
  filter(t1_acc == 1) |>
  group_by(sub, t1_level, t2_level) |>
  summarize(t2acc = mean(t2_acc), .groups="drop") |>
  pivot_wider(names_from = c(t1_level, t2_level),
              values_from = t2acc,
              names_prefix = "t2acc_") |>
  right_join(jasp_export,
             by=join_by(sub),
             unmatched="error",
             relationship="one-to-one") -> jasp_export
```

Add columns for demographics and the four survey sums

```{r addsurveysums}
surveyData |>
  select(sub, age, gender, ethnicity, highed,
         bsam=sumbsam, sias6=sumsias6,
         promisd=sumpromisd, promisa=sumpromisa) |>
  right_join(jasp_export,
             by=join_by(sub),
             unmatched="error",
             relationship="one-to-one") -> jasp_export
```

Create median splits for all columns

```{r mediansplits}
medians = jasp_export |>
  summarize(across(c(bsam, promisa, promisd, sias6),
                   ~median(.x, na.rm=TRUE)))
jasp_export |>
  mutate(anxiety_bsam = factor(ifelse(bsam > medians$bsam, "high", "low")),
         anxiety_pa = factor(ifelse(promisa > medians$promisa, "high", "low")),
         depression = factor(ifelse(promisd > medians$promisd, "high", "low")),
         anxiety_sias6 = factor(ifelse(sias6 > medians$sias6, "high", "low"))) |>
  relocate(sub, age, gender, ethnicity, highed,
           bsam, promisa, promisd, sias6,
           anxiety_bsam, anxiety_pa, anxiety_sias6, depression) -> jasp_export

jasp_export |> summary()
```

# Export Data

```{r, savefile}
write.csv(jasp_export, sprintf("RSVPLG02-JASP-%s.csv", Sys.Date()),
          row.names=FALSE)
```
