---
title: "Gather Data Files"
author: "David Fencsik"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
library(tidyverse)

this_file = "00-GatherData"
testmode = TRUE

# set variables to select from RSVP data file
variables = c(
  'exp',
  'ver',
  'mod.utc',
  'sub',
  'experimenter',
  'datetime',
  'blocktyp',
  'sess',
  'trial',
  'trial_type',
  'trial_time',
  't1_level',
  't2_level',
  't2_lag',
  'global_letters',
  'local_letters',
  't1_pos',
  't1',
  't1_corr',
  't1_resp',
  't1_acc',
  't2',
  't2_corr',
  't2_resp',
  't2_acc',
  't1_rt',
  't2_rt'
)
```

# Purpose

Gather all the individual RSVP data files and the survey data and save them as an R binary

# Load Data Files

## Combine Individual RSVP Data Files

Gather all the files in the data/ directory and all subdirectories together into one data frame, making sure that the headings match.

```{r loadrsvp1}
if (testmode) {
  dataDir = "data"
} else {
  dataDir = rstudioapi::selectDirectory(caption="Select RSVP Data Files Directory",
                                        label="Select",
                                        path=getwd())
}
if (is.null(dataDir)) {
  print("Canceled")
  knitr::knit_exit()
} else if (!dir.exists(dataDir)) {
  print(sprintf("Directory not found: %s", dataDir))
  knitr::knit_exit()
}
```

```{r, gatherdatacode, echo=FALSE}
GatherData = function(dir=NULL, columns=NULL) {

    if (is.null(dir) || !dir.exists(dir)) return (NULL)

    fileList = list.files(path=dir, pattern="\\.csv$", full.names=TRUE,
                          recursive=TRUE, ignore.case=TRUE, no..=TRUE)
    allData = NULL
    for (fileName in fileList) {
        df = read.csv(fileName)
        if (is.null(columns)) {
            columns = names(df)
        }
        if (!all(columns %in% names(df))) {
            warning("Skipping file due to missing columns: ", fileName)
            next
        }
        if (is.null(allData)) {
            allData = df[columns]
        } else {
            allData = rbind(allData, df[columns])
        }
    }
    return(allData)
}
```

```{r loadrsvp2}
rsvpData = GatherData(dataDir, variables)
sprintf("Gathered data files in %s together", dataDir)
sprintf("Data consists of %d rows and %d columns",
        dim(rsvpData)[1], dim(rsvpData)[2])
print("Columns:")
cat(sprintf("\"%s\"", paste(colnames(rsvpData), collapse="\", \"")))
```

## Load Survey Data

Load the survey CSV file

```{r loadsurvey1}
if (testmode) {
  surveyFile = "SurveyDataRaw.csv"
} else {
  surveyFile = rstudioapi::selectFile(caption="Select Survey Data File",
                                      label="Select",
                                      path=getwd(),
                                      filter="CSV Files (*.csv)",
                                      existing=TRUE)
}
if(is.null(surveyFile)) {
  print("Cancelled")
  knitr::knit_exit()
} else if (!file.exists(surveyFile)) {
  print(sprintf("File not found: %s", surveyFile))
  knitr::knit_exit()
}
```

```{r loadsurvey2}
surveyData = read.csv(surveyFile)
sprintf("Loaded survey data file %s", surveyFile)
sprintf("Data consists of %d rows and %d columns",
        dim(surveyData)[1], dim(surveyData)[2])
print("Columns:")
cat(sprintf("\"%s\"", paste(colnames(surveyData), collapse="\", \"")))
```

# Save Data

## Raw Task Data CSV File

Save the raw task data to a CSV file for sharing

```{r saversvp}
write.csv(rsvpData, sprintf("RSVPLG01-RawTaskData-%s.csv", Sys.Date()),
          row.names=FALSE)
```

## Combined Data R File

Save both the task data and the survey data to an R binary file

```{r savedata}
save(rsvpData, surveyData, file=sprintf("%s.RData", this_file))
```
