---
title: "RSVPLG01: Filter Data"
author: "David Fencsik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
this_file = "02-FilterData"
```

# Optionally Remove Participants

## Load data files

```{r, loaddata}
rsvpData = read.csv("rsvpData.csv")
surveyData = read.csv("surveyData.csv")
rsvpAggregate = read.csv("rsvpAggregate.csv")
```

## Number of Particants Before Filtering

```{r, preOutlierRemovalSummary}
print(length(unique(rsvpData$sub)))
print(length(unique(surveyData$sub)))
print(length(unique(rsvpAggregate$sub)))
```

## Outlier Removal

Remove SIAS6 outlier

```{r, outlierSIAS6}
filterSIAS6 = FALSE
if (filterSIAS6) {
  outlierSubs = c(135)
  rsvpAggregate = rsvpAggregate |>
    filter(!(sub %in% outlierSubs))
  rsvpData = rsvpData |>
    filter(!(sub %in% outlierSubs))
}
```

Remove PROMIS-A outlier(s)

```{r, outlierPROMISA}
filterPROMISA = FALSE
if (filterPROMISA) {
  outlierSubs = c(237)
  rsvpAggregate = rsvpAggregate |>
    filter(!(sub %in% outlierSubs))
  rsvpData = rsvpData |>
    filter(!(sub %in% outlierSubs))
}
```

Remove PROMIS-D outlier(s)

```{r, outlierPROMISD}
filterPROMISD = FALSE
if (filterPROMISD) {
  outlierSubs = c(244)
  rsvpAggregate = rsvpAggregate |>
    filter(!(sub %in% outlierSubs))
  rsvpData = rsvpData |>
    filter(!(sub %in% outlierSubs))
}
```

Remove subjects w/ fewer than 5 correct T1 responses in any cell

```{r, outliert1accbasic}
filtert1acc = FALSE
if (filtert1acc) {
  outlierSubs = c(109, 127, 132, 149, 162, 167, 180, 184, 187, 189,
                  194, 212, 217, 245, 248, 251, 255, 260, 261, 269,
                  277, 302, 304, 308, 312, 316, 325)
  rsvpAggregate = rsvpAggregate |>
    filter(!(sub %in% outlierSubs))
  rsvpData = rsvpData |>
    filter(!(sub %in% outlierSubs))
}
```

Remove subjects below the median T1 accuracy

```{r outliert1accextreme}
filtert1accextreme = FALSE
if (filtert1accextreme) {
  x = rsvpAggregate |>
    group_by(sub) |>
    summarize(t1acc = mean(t1acc), .groups="drop")
  t1acc_median = median(x$t1acc)
  subToKeep = x$sub[x$t1acc >= t1acc_median]
  rsvpAggregate = rsvpAggregate |>
    filter(sub %in% subToKeep)
  rsvpData = rsvpData |>
    filter(sub %in% subToKeep)
  surveyData = surveyData |>
    filter(sub %in% subToKeep)
}
```

Include just the participants run by Amanda's RAs in 2023--24. These are participants numbered in the 200s; David's students ran participants numbered in the 100s or 300s.

```{r justamanda}
filterjustamanda = FALSE
if (filterjustamanda) {
  rsvpAggregate = rsvpAggregate |>
    filter(sub >= 200 & sub <= 299)
  rsvpData = rsvpData |>
    filter(sub >= 200 & sub <= 299)
  surveyData = surveyData |>
    filter(sub >= 200 & sub <= 299)
}
```

Include just the participants run by David's 491C students in 2023--24 (see above). Note that only one of these two should be TRUE, otherwise we'll remove everyone.

```{r justdavid}
filterjustdavid = FALSE
if (filterjustdavid) {
  rsvpAggregate = rsvpAggregate |>
    filter(sub < 200 | sub > 299)
  rsvpData = rsvpData |>
    filter(sub < 200 | sub > 299)
  surveyData = surveyData |>
    filter(sub < 200 | sub > 299)
}
```

## Number of Particants After Filtering

```{r, postOutlierRemovalSummary}
print(length(unique(rsvpData$sub)))
print(length(unique(surveyData$sub)))
print(length(unique(rsvpAggregate$sub)))
```

## Save data

```{r, savedata}
save(rsvpData, surveyData, rsvpAggregate, file=sprintf("%s.RData", this_file))
```
