---
title: "RSVPLG01: Signal Detection Theory Analyses"
author: "David Fencsik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ez)
this_file = "07-SDT"
data_file = "02-FilterData.RData"
```

Analyze data based on signal detection theory, calculating d' and criterion

# Load data files

```{r, loaddata1}
if(!file.exists(data_file)) {
  sprintf("Cannot find file %s", data_file)
  knitr::knit_exit()
}
```

```{r, loaddata2}
load(data_file)
```

# Define function to calculate d'

```{r, dprimefunction}
ComputeDprime <- function (nhits, nfa, npos, nneg, correction=0.5) {
    ## correct HR for {0,1}
    adjHR <- nhits
    index <- adjHR == 0
    if (any(index)) adjHR[index] <- 0.5
    index <- adjHR == npos
    if (any(index)) adjHR[index] <- npos[index] - 0.5
    adjHR <- adjHR / npos

    ## correct FA for {0,1}
    adjFA <- nfa
    index <- adjFA == 0
    if (any(index)) adjFA[index] <- 0.5
    index <- adjFA == nneg
    if (any(index)) adjFA[index] <- nneg[index] - 0.5
    adjFA <- adjFA / nneg

    ## compute d' and criterion
    dprime <- qnorm(adjHR) - qnorm(adjFA)
    crit <- -0.5 * (qnorm(adjHR) + qnorm(adjFA))

    ## compute CI around d'
    phiFA <- 1 / sqrt(2 * pi) * exp(-0.5 * qnorm(adjFA))
    phiHR <- 1 / sqrt(2 * pi) * exp(-0.5 * qnorm(adjHR))
    ci <- 1.96 * sqrt(adjHR * (1 - adjHR) / npos / (phiHR^2) +
                      adjFA * (1 - adjFA) / nneg / (phiFA^2))

    return(data.frame(dprime=dprime, criterion=crit, ci=ci))
}
```

# Setup Data

## Compute counts

```{r, computecounts}
dt = rsvpData |>
  filter(trial_type == "exp",
         t1_acc == 1,
         t2_acc >= 0) |>
  mutate(t2 = if_else(t2 == "X", "present", "absent")) |>
  group_by(sub, lag, t1_level, t2_level, t2) |>
  summarize(ncor = sum(t2_acc),
            count = n(),
            .groups = "drop")
```

## Separate Data

Separate present and absent trials, then duplicate the absent trials so there's a set for both 300 and 750 ms lags.

```{r, separatetypes}
dt_present = dt |>
  filter(t2 == "present")

dt_absent_300 = dt |>
  filter(t2 == "absent") |>
  mutate(lag = 300)

dt_absent_750 = dt_absent_300 |>
  mutate(lag = 750)
```

## Recombine Data

```{r, combinedata}
# combined data
dt = bind_rows(dt_present, dt_absent_300, dt_absent_750)
```

## Compute D'

```{r, computedprime}
dt = dt |>
  pivot_wider(names_from=t2,
              values_from=c(ncor, count)) |>
  drop_na(ends_with(c("absent", "present"))) |>
  mutate(nerr_absent = count_absent - ncor_absent) |>
  select(sub, t1_level, t2_level, lag,
         n_hits=ncor_present, n_present=count_present,
         n_fa=nerr_absent, n_absent=count_absent) |>
  nest(data = c(n_hits, n_present, n_fa, n_absent)) |>
  mutate(out = map(data, ~ComputeDprime(.$n_hits, .$n_fa, .$n_present, .$n_absent))) |>
  unnest(cols=c(data, out)) |>
  select(!(c(n_hits, n_present, n_fa, n_absent)))
```

## Add in Demographics and Surveys

```{r, addothervars}
dt = rsvpAggregate |>
  filter(lag > 0) |>
  select(sub, t1_level, t2_level, lag,
         age, gender,
         bsam, sias6, promisa, promisd) |>
  full_join(dt) |>
  drop_na(bsam:ci)
```

## Filter

Remove participants who are missing a d' in any condition

```{r, filterlowdprime}
sub_to_keep = dt |>
  mutate(t1l=substr(t1_level, 1, 1),
         t2l=substr(t2_level, 1, 1),
         var=paste0(t1l, t2l, lag)) |>
  pivot_wider(names_from=var,
              values_from=dprime,
              id_cols=sub) |>
  drop_na() |>
  select(sub)

sub_to_keep = sub_to_keep$sub

dt = dt |>
  filter(sub %in% sub_to_keep)
```

# Median split on Promis-A

```{r, mediansplit}
x = dt |>
  distinct(sub, promisa)

dt = dt |>
  mutate(anxiety = factor(ifelse(promisa > median(x$promisa), "high", "low")))
```

# Plots

```{r, plotdprime}
dt |>
  mutate(scope=factor(paste(t1_level, t2_level, sep="-"))) |>
  group_by(scope, lag, anxiety) |>
  summarize(dprime=mean(dprime),
            .groups="drop") |>  
  ggplot(aes(x=lag, y=dprime, color=anxiety)) +
  geom_line(linewidth=1.25) +
  facet_wrap(~scope) +
  theme_classic()
```

# ANOVAs

```{r, anovas}
# get the 4 combos of scope
dt = dt |>
  mutate(scope=paste(t1_level, t2_level, sep="-"),
         sub=factor(sub),
         lag=factor(lag),
         anxiety=factor(anxiety))

scope = sort(unique(dt$scope))

for (s in scope) {
  cat("ANOVA on scope", s, "\n")
  # print(summary(dt[dt$scope == s, ]))
  dt |>
    filter(scope == s) |>
    ezANOVA(dv=dprime,
                wid=sub,
                within=lag,
                between=anxiety,
                type=3,
                detailed=FALSE) |>
    print()
}
```
