---
title: "Prepare and Analyze Data for Spring 2024 PSYC 300"
author: "David Fencsik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
library(viridis)
library(rstatix)
this_file = "91-AnalysesFor2024SP-300"
source_file = "01-CleanData"
dataFile = sprintf("%s.RData", source_file)
```

# Load data files

```{r, loaddata1}
if(!file.exists(dataFile)) {
  sprintf("Cannot find file %s", dataFile)
  knitr::knit_exit()
}
```

```{r, loaddata2}
load(dataFile)
```

# Simplify Data

Extract a simplified data set that removes the local/global distinction

```{r, extractdata}
x = rsvpAggregate |>
  filter(t2_lag > 0) |>
  group_by(sub, gender, age, t2_lag) |>
  summarize(t1acc = mean(t1acc),
            t2acc = mean(t2acc),
            t2t1acc = mean(t2t1acc),
            promisa = mean(promisa),
            promisd = mean(promisd),
            sias6 = mean(sias6),
            bsam = mean(bsam),
            .groups = "drop") |>
  drop_na(t2t1acc)
medians = x |>
  distinct(sub, .keep_all = TRUE) |>
  summarize(across(c(promisa, promisd), ~median(.x, na.rm=TRUE)))
x = x |>
  mutate(anxiety    = ifelse(promisa > medians$promisa, "high", "low"),
         depression = ifelse(promisd > medians$promisd, "high", "low"))
```

# Analysis

Conduct an ANOVA on the effects of T2 lag and anxiety on T2|T1 accuracy.

```{r, anova}
x |>
  anova_test(t2t1acc ~ t2_lag*anxiety +
             Error(sub/(t2_lag)), effect.size="pes")
```

Plot T2|T1 accuracy as a function of PROMIS-A.

```{r, plotpromisa}
x |>
  ggplot(aes(x=promisa, y=t2t1acc, color=factor(t2_lag), group=t2_lag)) +
  geom_point() +
  geom_smooth(method="lm", formula=y~x, se=FALSE)
```

Plot T2|T1 accuracy as a function of PROMIS-D.

```{r, plotpromisd}
x |>
  ggplot(aes(x=promisd, y=t2t1acc, color=factor(t2_lag), group=t2_lag)) +
  geom_point() +
  geom_smooth(method="lm", formula=y~x, se=FALSE)
```

Plot T2|T1 accuracy as a function of BSAM.

```{r, plotbsam}
x |>
  ggplot(aes(x=bsam, y=t2t1acc, color=factor(t2_lag), group=t2_lag)) +
  geom_point() +
  geom_smooth(method="lm", formula=y~x, se=FALSE)
```

Plot T2|T1 accuracy as a function of SIAS6.

```{r, plotsias6}
x |>
  ggplot(aes(x=sias6, y=t2t1acc, color=factor(t2_lag), group=t2_lag)) +
  geom_point() +
  geom_smooth(method="lm", formula=y~x, se=FALSE)
```

# Data Export

Export a wide data set for import into JASP

```{r, exportJASP}
jasp_export = x |>
  select(!t2acc) |>
  rename(t2acc = t2t1acc) |>
  pivot_wider(names_from=t2_lag,
              values_from=c(t1acc, t2acc),
              id_cols=c("sub", "gender", "age",
                        "bsam", "sias6", "promisa", "promisd",
                        "anxiety", "depression")) |>
  mutate(anxiety = factor(anxiety),
         depression = factor(depression))
summary(jasp_export)
write.csv(jasp_export, sprintf("RSVPLG01-JASP-%s.csv", Sys.Date()),
          row.names=FALSE)
```
