---
title: "Prepare and Analyze Data for Classes"
author: "David Fencsik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
library(viridis)
library(rstatix)
```

# Load data files

```{r, loaddata}
rsvpAggregate = read.csv("rsvpAggregate.csv")
```

# Create Wide Data Set

Generate a wide data set for import into JASP that has columns for T2 accuracy for each of the 8 cells in the factorial table of T1 level (local, global), T2 level (local, global), and T2 lag (300, 750)

```{r, generate1}
jasp_export = rsvpAggregate |>
  select(!t2acc) |>
  rename(t2acc = t2t1acc) |>
  filter(t2_lag > 0) |>
  mutate(t1_level = ifelse(t1_level == "global", "G", "L"),
         t2_level = ifelse(t2_level == "global", "G", "L"),
         t2_lag = t2_lag * 150,
         iv_names = paste("t2acc", t1_level, t2_level, t2_lag, sep="")) |>
  pivot_wider(names_from=iv_names,
              values_from=t2acc,
              id_cols=c("sub", "gender", "age",
                        "bsam", "promisa", "promisd", "sias6")) |>
  drop_na(starts_with("t2acc"))
```

Create median splits for all columns

```{r, generate2}
medians = jasp_export |>
  distinct(sub, .keep_all = TRUE) |>
  summarize(across(c(bsam, promisa, promisd, sias6),
                   ~median(.x, na.rm=TRUE)))
jasp_export = jasp_export |>
  mutate(anxiety_bsam = factor(ifelse(bsam > medians$bsam, "high", "low")),
         anxiety_pa = factor(ifelse(promisa > medians$promisa, "high", "low")),
         depression = factor(ifelse(promisd > medians$promisd, "high", "low")),
         anxiety_sias6 = factor(ifelse(sias6 > medians$sias6, "high", "low"))) |>
  relocate(sub, gender, age,
           bsam, promisa, promisd, sias6,
           anxiety_bsam, anxiety_pa, anxiety_sias6, depression)
```

Add columns with in T2 accuracy collapsing across all the local/global columns

```{r, generate3}
jasp_export = jasp_export |>
  rowwise() |>
  mutate(t2acc300 = mean(c_across(ends_with("300"))),
         t2acc750 = mean(c_across(ends_with("750")))) |>
  relocate(t2acc300, t2acc750, .after=depression) |>
  ungroup()
summary(jasp_export)
```

Add columns comparing switch (GL and LG) to no-switch trials for each lag

```{r generate4}
jasp_export = jasp_export |>
  rowwise() |>
  mutate(t2acc300noswitch =
           mean(c_across(ends_with("300") & contains(c("GG", "LL")))),
         t2acc300switch =
           mean(c_across(ends_with("300") & contains(c("GL", "LG")))),
         t2acc750noswitch =
           mean(c_across(ends_with("750") & contains(c("GG", "LL")))),
         t2acc750switch =
           mean(c_across(ends_with("300") & contains(c("GL", "LG")))))
```

Generate and add columns with T1 accuracy at each level of T1, collapsing across T2 variables`

```{r, generate5}
t1 = rsvpAggregate |>
  mutate(t1_level = ifelse(t1_level == "global", "G", "L")) |>
  group_by(sub, t1_level) |>
  summarize(t1acc=mean(t1acc), .groups="drop") |>
  pivot_wider(names_from=t1_level,
              values_from=t1acc,
              id_cols=sub) |>
  drop_na()
jasp_export$t1accG = rep(NA, nrow(jasp_export))
jasp_export$t1accL = rep(NA, nrow(jasp_export))
for (sub in jasp_export$sub) {
  if (sub %in% t1$sub) {
    i1 = jasp_export$sub == sub
    i2 = t1$sub == sub
    jasp_export$t1accG[i1] = t1$G[i2]
    jasp_export$t1accL[i1] = t1$L[i2]
  }
}
summary(jasp_export)
```

# Export Data

```{r, savefile}
write.csv(jasp_export, sprintf("RSVPLG01-JASP-%s.csv", Sys.Date()),
          row.names=FALSE)
```
