---
title: "RSVPLG03: Gather Data Files"
author: "David Fencsik"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
library(tidyverse)

this_file = "00-GatherData"
```

# Purpose

Gather all the individual RSVP data files and the survey data and save them as an R binary

# Load Data Files

## Combine Individual RSVP Data Files

Gather all the files in the data/ directory and all subdirectories together into one data frame, making sure that the headings match.

```{r, gatherdatacode, echo=FALSE}
GatherData = function(dir=NULL, columns=NULL) {

    if (is.null(dir) || !dir.exists(dir)) return (NULL)

    fileList = list.files(path=dir, pattern="\\.csv$", full.names=TRUE,
                          recursive=TRUE, ignore.case=TRUE, no..=TRUE)
    allData = NULL
    for (fileName in fileList) {
        df = read.csv(fileName)
        if (is.null(columns)) {
            columns = names(df)
        }
        if (!all(columns %in% names(df))) {
            warning("Skipping file due to missing columns: ", fileName)
            next
        }
        if (is.null(allData)) {
            allData = df[columns]
        } else {
            allData = rbind(allData, df[columns])
        }
    }
    return(allData)
}
```

```{r loadrsvp1}
dataDir = "data"
rsvpData = GatherData(dataDir)
sprintf("Gathered data files in %s together", dataDir)
sprintf("Data consists of %d rows and %d columns",
        dim(rsvpData)[1], dim(rsvpData)[2])
print("Columns:")
cat(sprintf("\"%s\"", paste(colnames(rsvpData), collapse="\", \"")))
```

## Load Survey Data

Load the survey CSV file

```{r loadsurvey1}
surveyFile = "SurveyDataRaw.csv"
surveyData = read.csv(surveyFile)
sprintf("Loaded survey data file %s", surveyFile)
sprintf("Data consists of %d rows and %d columns",
        dim(surveyData)[1], dim(surveyData)[2])
print("Columns:")
cat(sprintf("\"%s\"", paste(colnames(surveyData), collapse="\", \"")))
```

# Save Data

## Raw Task Data CSV File

Save the raw task data to a CSV file for sharing

```{r saversvp}
write.csv(rsvpData, sprintf("RSVPLG03-RawTaskData-%s.csv", Sys.Date()),
          row.names=FALSE)
```

## Combined Data R File

Save both the task data and the survey data to an R binary file

```{r savedata}
save(rsvpData, surveyData, file=sprintf("%s.RData", this_file))
```
