---
title: "RSVPLG03: Switch-Cost Analyses"
author: "David Fencsik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggpubr)
library(rstatix)
theme_set(theme_classic())
this_file = "08-Switching"
source_file = "02-FilterData"
dataFile = sprintf("%s.RData", source_file)
```

# Summary

The purpose of this analysis is to look at switch costs, starting with whether there is any evidence of switching between scopes.

# Load data files

```{r, loaddata1, include=FALSE}
load(dataFile)
```

# Local vs. Global Trials

T1 accuracy as a function of level

```{r T1accbylevel, fig.height=4, fig.width=4}
rsvpData |>
  filter(blocktype == "Experiment",
         trialtype == "main",
         t1_acc >= 0,
         t2_acc >= 0) |>
  group_by(sub, t1_level) |>
  summarize(t1acc = mean(t1_acc),
            .groups="drop") -> x

x |>
  ggplot(aes(x=t1_level, y=t1acc)) +
  geom_violin(quantile.linetype=1, quantile.color="gray") +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  geom_hline(yintercept=0.5, linetype=2, color="gray") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="T1 Level", y="T1 Accuracy") +
  scale_x_discrete(labels=c("global"="Global", "local"="Local"))
```

T2|T1 accuracy as a function of level

```{r T2accbylevel, fig.height=4, fig.width=4}
rsvpData |>
  filter(blocktype == "Experiment",
         trialtype == "main",
         t1_acc == 1,
         t2_acc >= 0) |>
  group_by(sub, t2_level) |>
  summarize(t2t1acc = mean(t2_acc),
            .groups="drop") -> x

x |>
  ggplot(aes(x=t2_level, y=t2t1acc)) +
  geom_violin(quantile.linetype=1, quantile.color="gray") +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  geom_hline(yintercept=0.5, linetype=2, color="gray") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="T2 Level", y="T2|T1 Accuracy") +
  scale_x_discrete(labels=c("global"="Global", "local"="Local"))
```

# Switch vs. Repeat Trials

Prepare data

```{r filterdata}
rsvpData |>
  filter(blocktype == "Experiment",
         trialtype == "main",
         t1_acc >= 0,
         t2_acc >= 0) |>
  mutate(switch_type=case_when(
           t1_level == t2_level ~ "noswitch",
           t1_level != t2_level ~ "switch",
           .default = NA)) |>
  filter(t1_acc == 1) |>
  group_by(sub, t2_level, switch_type) |>
  summarize(t2t1acc = mean(t2_acc),
            .groups="drop") -> x

summary(x)
```

Generate violin plots of overall repeating and switching

```{r ViolinPlotsBySwitch, fig.height=4, fig.width=4}
x |>
  group_by(sub, switch_type) |>
  summarize(t2t1acc = mean(t2t1acc), .groups="drop") |>
  ggplot(aes(x=switch_type, y=t2t1acc)) +
  geom_violin(quantile.linetype=1, quantile.color="gray") +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  geom_hline(yintercept=0.5, linetype=2, color="gray") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="Switch Condition", y="T2|T1 Accuracy") +
  scale_x_discrete(labels=c("noswitch"="No Switch", "switch"="Switch"))
```

Generate violin plots of switch costs

```{r ViolinPlotsSwitchCost, fig.height=4, fig.width=3}
x |>
  group_by(sub, switch_type) |>
  summarize(t2t1acc = mean(t2t1acc), .groups="drop") |>
  pivot_wider(names_from=switch_type,
              values_from=t2t1acc) |>
  mutate(cost = noswitch - switch) -> switch.cost

switch.cost |>
  ggplot(aes(x="", y=cost)) +
  geom_violin(quantile.linetype=1, quantile.color="gray") +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  geom_hline(yintercept=0, linetype=2, color="gray") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="", y="No Switch - Switch")
```

Generate violin plots of overall repeating and switching separated by level

```{r ViolinPlotsByLevelAndSwitch, fig.height=4, fig.width=8}
x |>
  ggplot(aes(x=switch_type, y=t2t1acc)) +
  geom_violin(quantile.linetype=1, quantile.color="gray") +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  geom_hline(yintercept=0.5, linetype=2, color="gray") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="Switch Condition", y="T2|T1 Accuracy") +
  scale_x_discrete(labels=c("noswitch"="No Switch", "switch"="Switch")) +
  facet_wrap(~t2_level, labeller=as_labeller(c(global="T2 Global", local="T2 Local")))
```

Generate violin plots of switch costs separated by level

```{r ViolinPlotsSwitchCostByLevel, fig.height=4, fig.width=4}
x |>
  pivot_wider(names_from=switch_type,
              values_from=t2t1acc) |>
  mutate(cost = noswitch - switch) -> switch.cost

switch.cost |>
  ggplot(aes(x=t2_level, y=cost)) +
  geom_violin(quantile.linetype=1, quantile.color="gray") +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  geom_hline(yintercept=0, linetype=2, color="gray") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="T2 Level", y="No Switch - Switch") +
  scale_x_discrete(labels=c("global"="Global", "local"="Local"))
```

Add survey totals to switch.cost data

```{r AddSurveysToSwitchCost}
surveyData |>
  select(sub, bsam, cesdr, sias6, sticsa) |>
  right_join(switch.cost,
             by=join_by(sub),
             unmatched="error",
             relationship="one-to-many") -> switch.cost
```

Plot local and global switch cost as a function of STICSA total

```{r ScatterplotSwitchCostByLevelAndAnxiety}
switch.cost |>
  mutate(t2_level = factor(t2_level, levels=c("global", "local"),
                           labels=c("Global", "Local"))) |>
  ggplot(aes(x=sticsa, y=cost, color=t2_level)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  scale_color_manual(values=c("black","red")) +
  labs(x="Anxiety", y="No Switch - Switch (T2|T1 Accuracy)") +
  guides(color=guide_legend(title="T2 Level")) +
  theme(legend.position="inside",
        legend.position.inside=c(0.9, 0.9))
```

Linear fit of global data above. Runs a regression of the effect of trait anxiety (STICSA) on the difference score for just T2 global.

```{r RegressionSwitchCostBySTICSAforGlobal}
fit = lm(cost ~ sticsa, data=switch.cost, subset=(t2_level == "global"))
summary(fit)
confint(fit, 'sticsa', level=0.95)
```

Same as above but for T2 local.

```{r RegressionSwitchCostBySTICSAforLocal}
fit = lm(cost ~ sticsa, data=switch.cost, subset=(t2_level == "local"))
summary(fit)
confint(fit, 'sticsa', level=0.95)
```
