---
title: 'RSVPLG03: Attentional Blink'
author: "David Fencsik, Amanda Morrison"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(rstatix)
library(knitr)
theme_set(theme_classic())
this_file = "11-Blink"
source_file = "02-FilterData"
dataFile = sprintf("%s.RData", source_file)
load(dataFile)
```

# Setup Data

Prepare AB data

```{r ExtractABData}
# extract T1 accuracy
rsvpData |>
  filter(blocktype == "Experiment",
         trialtype == "main",
         t1_acc >= 0,
         t2_acc >= 0) |>
  group_by(sub, lag) |>
  summarize(t1acc = mean(t1_acc),
            .groups = "drop") -> ab.data

# extract T2|T1 accuracy and add to data
rsvpData |>
  filter(blocktype == "Experiment",
         trialtype == "main",
         t1_acc == 1,
         t2_acc >= 0) |>
  group_by(sub, lag) |>
  summarize(t2t1acc = mean(t2_acc),
            .groups = "drop") |>
  right_join(ab.data,
             by=join_by(sub, lag),
             unmatched="error",
             relationship="one-to-one") -> ab.data

# add survey data
surveyData |>
  select(sub, bsam, cesdr, sias6, sticsa) |>
  right_join(ab.data,
             by=join_by(sub),
             unmatched="error",
             relationship="one-to-many") -> ab.data
```

# Attentional Blink

Here's a plot based on our data showing T1 and T2|T1 accuracy as a function of lag. As expected, T1 accuracy doesn't change much with lag while T2|T1 accuracy is lower at lags of 300 ms compared to lags of 750 ms. Error bars are 95% confidence intervals based on the difference in each line.

```{r BasicABPlot, fig.height=4, fig.width=4}
# extract 95% CI ranges for error bars
ab.data |>
  t_test(t1acc ~ lag, paired=TRUE, detailed=TRUE) |>
  select(conf.low, conf.high) |>
  mutate(ci.range = abs(conf.low - conf.high)) |>
  pull(ci.range) -> ci.t1

ab.data |>
  t_test(t2t1acc ~ lag, paired=TRUE, detailed=TRUE) |>
  select(conf.low, conf.high) |>
  mutate(ci.range = abs(conf.low - conf.high)) |>
  pull(ci.range) -> ci.t2

ab.data |>
  group_by(lag) |>
  summarize(t1acc = mean(t1acc),
            t2t1acc = mean(t2t1acc),
            .groups="drop") |>
  pivot_longer(cols = ends_with("acc"),
               names_to = "target",
               values_to = "acc") |>
  mutate(lag = factor(lag),
         lower=case_when(
           target == "t1acc" ~ acc - ci.t1,
           target == "t2t1acc" ~ acc - ci.t2,
           .default = 0
         ),
         upper=case_when(
           target == "t1acc" ~ acc + ci.t1,
           target == "t2t1acc" ~ acc + ci.t2,
           .default = 0
         ),
         target = factor(target, levels=c("t1acc", "t2t1acc"),
                         labels = c("T1", "T2|T1"))) |>
  ggplot(aes(x=lag, y=acc, group=target, color=target,
             ymin=lower, ymax=upper)) +
  geom_line(linewidth=1.25) +
  geom_errorbar(width = 0.1) +
  geom_point(size=3, shape=21, fill="white", stroke=2) +
  scale_color_manual(values=c("black","red")) +
  coord_cartesian(ylim=c(0.5, 1.0)) +
  labs(x="Lag (ms)", y="Proportion correct") +
  theme(legend.position="inside",
        legend.position.inside=c(0.8, 0.2),
        legend.title=element_blank())
```

Here's a violin plot showing the distributions for each of the points that went into the above line graph:

```{r ViolinPlotsByTargetAndLag, fig.height=4, fig.width=8}
ab.data |>
  pivot_longer(cols = ends_with("acc"),
               names_to = "target",
               values_to = "acc") |>
  mutate(lag = factor(lag)) |>
  ggplot(aes(x=lag, y=acc)) +
  geom_violin(quantile.linetype=1, quantile.color="gray") +
  geom_jitter(width=.15, height=0, color="black", shape="circle open") +
  geom_hline(yintercept=0.5, linetype=2, color="gray") +
  stat_summary(fun=mean, geom="point", color="red", shape="cross", size=5) +
  labs(x="Lag (ms)", y="Accuracy") +
  facet_wrap(~target, labeller=as_labeller(c(t1acc="T1", t2t1acc="T2|T1")))
```

## Analyses

A *t*-test of the effects of lag on T1 accuracy, starting with a QQ-plot and a Shapiro-Wilk test to evaluate normality. T1 accuracies look non-normal (as of 1/1/2026) but we don't really care much about them.

```{r ttestT1, fig.height=4, fig.width=4}
ab.data |>
  ggplot(aes(sample=t1acc, color=factor(lag))) +
  stat_qq() +
  stat_qq_line()

ab.data |>
  group_by(lag) |>
  shapiro_test(t1acc)

ab.data |>
  t_test(t1acc ~ lag, paired=TRUE)
```

Same for T2|T1 accuracy. These seem more normal.

```{r ttestT2, fig.height=4, fig.width=4}
ab.data |>
  ggplot(aes(sample=t2t1acc, color=factor(lag))) +
  stat_qq() +
  stat_qq_line()

ab.data |>
  group_by(lag) |>
  shapiro_test(t2t1acc)

ab.data |>
  t_test(t2t1acc ~ lag, paired=TRUE)
```



# Effects of Anxiety

Here's a scatterplot of T2|T1 accuracy as a function of trait anxiety, separating out each of the two lags, with the best fit line for each.

```{r ScatterplotT2AccByLagAndSTICSAFaceted, fig.height=4, fig.width=10}
ab.data |>
  mutate(lag = factor(lag)) |>
  ggplot(aes(x=sticsa, y=t2t1acc)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  labs(x="Anxiety", y="T2|T1 Accuracy") +
  guides(color=guide_legend(title="Lag (ms)")) +
  facet_wrap(~lag)
```

Here's the same data but with the two lags on the same graph.

```{r ScatterplotT2AccByLagAndSTICSA, fig.height=4, fig.width=6}
ab.data |>
  mutate(lag = factor(lag)) |>
  ggplot(aes(x=sticsa, y=t2t1acc, color=lag)) +
  geom_point(size=1) +
  geom_smooth(method="lm", formula=y~x, se=FALSE) +
  scale_color_manual(values=c("black","red")) +
  labs(x="Anxiety", y="T2|T1 Accuracy") +
  guides(color=guide_legend(title="Lag (ms)")) +
  theme(legend.position="inside",
        legend.position.inside=c(0.9, 0.15))
```

Linear fit of lag 300 data above. Runs a regression of the effect of STICSA on T2|T1 accuracy at lag 300

```{r RegressionSwitchCostBySTICSAlag300}
fit = lm(t2t1acc ~ sticsa, data=ab.data, subset=(lag == 300))
summary(fit)
confint(fit, 'sticsa', level=0.95)
```

Same as above but for lag 750.

```{r RegressionSwitchCostBySTICSAlag750}
fit = lm(t2t1acc ~ sticsa, data=ab.data, subset=(lag == 750))
summary(fit)
confint(fit, 'sticsa', level=0.95)
```

First rows of the data file:
```{r PreviewData, results='asis'}
ab.data |>
  head(10) |>
  kable()
```
